---
title: "RM HW #4 RUSSMAN"
output: html_notebook
---

The cycle of crime theory asserts that overly harsh sentencing leads to more crime

This is an interesting design and seems to touch on mechanisms behind the theory of interest. As researchers we cannot manipulate length of sentence, however we view sentencing as a causal shock and look at downstream outcomes. This implies that sentencing has some variability. Depending on who you ask this may or may not be true. Judges and juries are human and prone to bias and influence so there is variability in sentencing, however the variability may pertain to race, age, gender, personality/likeability and other factors and these factors would need to be controlled for. The sentencing effects jail time which in turn may have a causal influence on repeat offenses, with the theory that higher jail time leads to higher likelihood of recidivism. 

```{r}
library(tidyverse)
library(cobalt)
library(stargazer)
library(AER)
rm(list=ls())
setwd("~/Documents/2.Research Methods/Week 11/assignment 4")
data=read.csv("crime-iv.csv")
View(data)

```


```{r}
#balance test
balance_table = bal.tab(data[,c("Severity.Of.Crime","Months.In.Jail")], treat=data$Republican.Judge, disp = c("means", "sds"))
balance_table= balance_table$Balance[,1:6]
balance_table$P_Value=c(t.test(data$Severity.Of.Crime~data$Republican.Judge)$p.value,t.test(data$Months.In.Jail~data$Republican.Judge)$p.value)
balance_table

```

groups are statistically significantly different along severity of crime and month in jail by judge, does not appear randomly assigned. The republican judges are assigned to the less severe crimes and doling out more severe punishments, with the mean difference of 3 months in jail. If we ignore the fact that judges may not be randomly assigned, or if we assume randommness in the severity of crime designation then republican judges as an instrument variable does result in statistically different months in jail a.k.a treatments lead to differing compliance

in the first stage OLS we will use the republican judge as the instrument variable (assuming random assignment) and predict the jail sentence, while controlling for the severity of the crime

```{r}
#first stage OLS
model1=lm(data=data, Months.In.Jail ~ Republican.Judge + Severity.Of.Crime)
summary(model1)

stargazer(model1, type='text', title= 'First Stage OLS for Republican Judge Instrument Variable on Jail Sentence', out="assignment4-first stage.txt", digits=2, covariate.labels=c("Republican Judge","Severity of Crime", "Intercept"), dep.var.labels=c("Months in Sentence"))

```

the interpretation of the resulting model is that individuals with a republican judge receive an additional 3.22 months of sentence than those without a republican judge

```{r}
#second stage OLS/ reduced form

model2=lm(data=data, Recidivates ~ Republican.Judge + Severity.Of.Crime)
summary(model2)

stargazer(model2, type='text', title= 'Second Stage OLS for Jail Sentence on Recidivism', out="assignment4-second stage.txt", digits=2, covariate.labels=c("Republican Judge","Severity of Crime", "Intercept"), dep.var.labels=c("Recidivism"))

```

```{r}
ratio = model2$coefficients[2]/model1$coefficients[2]
ratio
```

The ratio of: Repeat offense by republican judge (.14) divided by months in jail by the IV of republican judge (3.22) is .044, signifying that on average, 4.4% of individuals sentenced by a republican judge commit another crime

```{r}
model3=lm(data=data, Recidivates ~ model1$fitted.values + Severity.Of.Crime)
summary(model3)

summary(aov(model3))
F_val=((summary(aov(model3))[[1]]$"Sum Sq"[1])+summary(aov(model3))[[1]]$"Sum Sq"[2])/2/(summary(aov(model3))[[1]]$"Sum Sq"[3]/summary(aov(model3))[[1]]$"Df"[3])
F_val

```

The F-value is highly significant F(2,4997)=366.2
The ratio method resulted in a value equal to the coefficent value of .044, signifying 4.4% of randomly assigned to republican judge repeat offense.

```{r}
iv_reg=ivreg(data=data, Recidivates ~ Months.In.Jail + Severity.Of.Crime | Republican.Judge + Severity.Of.Crime)
summary(iv_reg)
```

using an R function IV_reg i get the same results with a coefficient of .044. Should we be running a logit model instead of linear model though?

------------
In the research design above (using randomized judges), the always-takers are the criminals who are always going to get a longer sentence for their severity of crime, no matter the party affiliation of the judge. This could be because they are especially hostile in trial and certain biases towards longer sentencing work against their favor.

The never-takers are the criminals who are always going to get a shorter sentence no matter the judge's political affiliation. These are the kinder, more innocent seeming individuals where biases work in their favor to reduce their sentence length.

The compliers are the individuals who get a longer sentence with the republican judges. Perhaps these are democrats whom don't get along with the judge and clash in trial, resulting in a longer sentence

The defiers are the individuals who receive shorter sentences with the republican judges. Maybe these are republicans whom appeal to the judge in trial and whom the judges take pity on and shorten their sentences.

------------


We assume that the model is monotonic and that republican judges always give out longer sentences. There are no defiers in this case, in which the judges take more pity on and shorten the sentences for. This may not be true in real life as republican judges are still susceptible to human bias and will probably give shorter sentences to more attractive, innocent seeming individuals.

In this dataset the compliers are those that receive a longer sentence when the judges are republican than if they were democrat.

This analysis supports the cycle of crime hypothesis that harsher sentencing increases the likelihood of committing a repeat offense.
