---
title: "HW3 Research Methods 2 - Russman"
output: html_notebook
---

```{r}
library(cobalt)
library(tidyverse)
library(stargazer)

setwd("~/Documents/2.Research Methods/Week 10/Assignment 3")
data=read.csv("sports-and-education.csv")
View(data)
```


```{r}
#create balance table
balance_table = bal.tab(data[,c("Academic.Quality","Athletic.Quality","Near.Big.Market")], treat=data$Ranked.2017)
balance_table= balance_table$Balance[,1:2]
balance_table$P_Value=c(t.test(data$Academic.Quality~data$Ranked.2017)$p.value,t.test(data$Athletic.Quality~data$Ranked.2017)$p.value,t.test(data$Near.Big.Market~data$Ranked.2017)$p.value)
balance_table
```
treatment and control groups differ significantly on athletic quality and near big market

```{r}
model = lm(data=data, data$Ranked.2017 ~ data$Academic.Quality + data$Athletic.Quality + data$Near.Big.Market + data$Alumni.Donations.2018)
summary(model)

model2 = glm(data=data, family=binomial, data$Ranked.2017 ~ data$Academic.Quality + data$Athletic.Quality + data$Near.Big.Market + data$Alumni.Donations.2018)
summary(model2)

model3 = glm(data=data, family=binomial, data$Ranked.2017 ~ data$Athletic.Quality + data$Academic.Quality + data$Alumni.Donations.2018)
summary(model3)
```

removed near big market because the model had an R^2 of 1 and perfectly predicted the treatment. new model has R^2 near .36 and has some variability in the p-scores. Propensity score methods won't be effective in this case if all variables are included and the p-score perfectly matches the true treatment. There won't be a way to match or block very well without variability in the p-scores

```{r}
prop_score=data.frame(p_score = predict(model3, type= "response"), ranked = data$Ranked.2017)
data$propensity_score=predict(model3, type= "response")
prop_score
```

```{r}
ggplot(filter(data,data$Ranked.2017==1),aes(x = propensity_score, color = "blue")) + geom_histogram(binwidth = .02, fill = "black", alpha=.5, position = "identity") + geom_density() + theme_minimal()

ggplot(filter(data,data$Ranked.2017==0),aes(x = propensity_score, color = "green")) + geom_histogram(binwidth = .02, fill = "blue", alpha=.5, position = "identity") + geom_density() + theme_minimal()

```

```{r}
#adding blocks
data=data[order(data$propensity_score), ]

data$block[1:10]=1
data$block[10:20]=2
data$block[20:30]=3
data$block[30:40]=4
data$block[40:50]=5
data$block[50:60]=6
data$block[60:70]=7
data$block[70:80]=8
data$block[80:90]=9
data$block[90:100]=10

```

```{r}
model4=lm(data=data, data$Alumni.Donations.2018 ~ data$Athletic.Quality + data$College.Id + data$Near.Big.Market + data$Academic.Quality + data$block)
summary(model4)

stargazer(model4, type='text', title= 'Model for Alumni Donations with Ranked Propensity Score Blocking', out="assignment3.txt", digits=2, covariate.labels=c("Athletic Quality","College Fixed Effects", "Near Big Market","Academic Quality","Ranked Propensity Score Blocking", "Intercept"), dep.var.labels=c("Alumni Donations"),notes = c("The above table demonstrates", "the positive effect on alumni donations", "of a college being ranked top 50"))

model5=lm(data=data, data$Alumni.Donations.2018 ~ data$Athletic.Quality + data$College.Id + data$Near.Big.Market + data$Academic.Quality + data$Ranked.2017)
summary(model5)
```

the propensity score model estimates that alumni donations increase by $200K if the school is ranked (with removal of the near big market variable). While using the true ranking value from the data results in an estimate of $500K increase in alumni donations if ranked in the top 50. The Propensity
