---
title: "RM Assignment 2"
output: html_notebook
---


```{r}
library(tidyverse)
library(stargazer)
setwd("~/Documents/2.Research Methods/Week 9")
data=read.csv("vaping-ban-panel.csv")
View(data)
```

```{r}
#plot in GGplot
ggplot(data,aes(Year,Lung.Hospitalizations,group=State.Id,color=Vaping.Ban))+stat_summary(geom = 'line')+geom_vline(xintercept=2021)+theme_minimal()

data$Group = factor(if_else(data$State.Id<=23,1,0))
ggplot(data,aes(Year,Lung.Hospitalizations,group=Group,color=Vaping.Ban))+stat_summary(geom = 'line')+geom_vline(xintercept=2021)+theme_minimal()
```


```{r}
treatment = vector(length=length(table(data$Year)))
control = vector(length=length(table(data$Year)))
year = vector(length=length(table(data$Year)))
hosp_t=vector(length=length(table(data$Year)))
hosp_c=vector(length=length(table(data$Year)))

for (i in c(1:length(table(data$Year)))) {
  d=filter(data,data$Year==(i+2009) & data$State.Id <24)
  hosp_t[i]=mean(d$Lung.Hospitalizations)
  treatment[i]=mean(d$Vaping.Ban)
  d=filter(data,data$Year==(i+2009) & data$State.Id >=24)
  hosp_c[i]=mean(d$Lung.Hospitalizations)
  year[i]=2009+i
  control[i]=mean(d$Vaping.Ban)
}

data1=data.frame(cbind(year,hosp_t,hosp_c))

ggplot(data1, aes(x=year)) + geom_line(aes(y=hosp_t), color = "green") + geom_line(aes(y=hosp_c),color='red')

```

```{r}
#analyze parallel trends

data2=filter(data,data$Year<2021)

ggplot(data2,aes(Year,Lung.Hospitalizations,group=Group,color=Vaping.Ban))+stat_summary(geom = 'line')+theme_minimal()

partrends=lm(data=data2,Lung.Hospitalizations ~ Year + Group + Year*Group)
summary(partrends)

stargazer(partrends, type='text', title= ' Table 1: Parallel Trends Analysis', out="ParTrends.txt", digits=2, covariate.labels=c("Year","Treatment States", "Treatment States by Year Interaction"), dep.var.labels=c("Number of Hospitalizations Per Year"))
```

```{r}
#run DID regression
didreg=lm(data=data, data$Lung.Hospitalizations ~ data$Year*data$Vaping.Ban)
summary(didreg)

stargazer(didreg, type='text', title= ' Table 2: Difference in Difference Analysis', out="DID.txt", digits=2, covariate.labels=c("Year","Vaping Ban Treatment", "Difference in Difference Estimator"), dep.var.labels=c("Number of Hospitalizations Per Year"))
```







There are 2 state level fixed effect (vaping ban and state). The vaping ban results in a significant change in hospitalizations within the states that receive it. States 1-23 are significantly different than states 24-50 (i.e. the states that are to receive the ban are already different than the states that won't receive the ban). Yes, we reject the hypothesis
that all state fixed effects are zero